Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2020-12-30 xrb2/paper.tex
Edited flame scaling paragraph in Section 3.2

We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms 
of temperature but differing rotation rates. {\color{blue}We saw in \citet{flame_wave1} that 
increasing the 
neutron star rotation rate reduces the horizontal lengthscale of the flame. An estimate of 
this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius may be 
thought of as the scale over which the balance between the Coriolis force and horizontal pressure 
gradient becomes important. It is approximately:
\begin{equation}
	\label{eqn:rossby}
	L_R \approx \frac{\sqrt{g H_0}}{\Omega},
\end{equation}
where $g$ is gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
neutron star rotation rate. Comparing Figure \ref{fig:time_series_enuc_500} and Figure 
\ref{fig:time_series_enuc_1000}, the horizontal extent of the $\tt{F1000}$ flame does appear to be 
reduced compared to the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:rossby}, increasing 
the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the greater 
confinement from the Coriolis force should reduce the horizontal extent of the flame by a similar 
factor. The Rossby radius is not a perfect measure of this horizontal lengthscale, however, and 
here the difference is not a factor of two. The simulations seem to follow the theoretical
scaling described in Equation \ref{eqn:rossby} more closely at higher rotation rates (1000 Hz and 
up), based on the results of \citet{flame_wave1}.}

@@

We found in \citet{flame_wave1} that the flame speed, $s$, obeys the following relation:
\begin{equation}
	\label{eqn:flame_speed_scaling}
	s \appropto L_R (\enucdot)^{\half} \propto \frac{(\enucdot)^{\half}}{\Omega},
\end{equation}
where $\enucdot$ is the specific nuclear energy generation rate and $L_R$ is the Rossby length 
described by Equation \ref{eqn:rossby}. 
This was consistent with the results of \citet{cavecchi:2013}, who noted that the flame speeds in 
their simulations scaled with the Rossby length. As seen in Figure~\ref{fig:flame_speeds_1}, there 
is no obvious inverse scaling of the flame speed with rotation rate in this set of runs. We 
observed earlier, however, that nuclear reactions progress more quickly at higher rotation 
rate. This results in a higher $\enucdot$ --- up to four to six times higher near the burning front 
after the flame ignites --- which may counteract the reduction in flame speed from the increased 
Coriolis confinement. We also observe that $\tt{F500}$ accelerates 
faster than $\tt{F1000}$, which appears to experience a small deceleration at early times. 
\MarginPar{KE: This is speculation on my part -- does someone else have clear idea why?} This 
disparity may be a direct result of the difference in Coriolis force. It could also be due to the 
500 Hz flame taking more time to develop fully and reach higher burning rates, resulting in an 
$\enucdot$ increase over time.

>>>>

We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms 
of temperature but differing rotation rates. {\color{blue}We saw in \citet{flame_wave1} that 
increasing the 
neutron star rotation rate reduces the horizontal lengthscale of the flame. An estimate of 
this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius may be 
thought of as the scale over which the balance between the Coriolis force and horizontal pressure 
gradient becomes important, and is approximately:
\begin{equation}
	\label{eqn:rossby}
	L_R \approx \frac{\sqrt{g H_0}}{\Omega},
\end{equation}
where $g$ is gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
neutron star rotation rate. Comparing Figure \ref{fig:time_series_enuc_500} and Figure 
\ref{fig:time_series_enuc_1000}, the horizontal extent of the $\tt{F1000}$ flame does appear to be 
reduced compared to the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:rossby}, increasing 
the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the greater 
confinement from the Coriolis force should reduce the horizontal extent of the flame by a similar 
factor. The Rossby radius is not a perfect measure of this horizontal lengthscale, however, and 
here the difference is not a factor of two. The simulations seem to follow the theoretical
scaling described in Equation \ref{eqn:rossby} more closely at higher rotation rates (1000 Hz and 
up), based on the results of \citet{flame_wave1}.}

@@

We found in \citet{flame_wave1} that the flame speed, $s$, obeys the following relation:
\begin{equation}
	\label{eqn:flame_speed_scaling}
	s \appropto L_R (\enucdot)^{\half} \propto \frac{(\enucdot)^{\half}}{\Omega},
\end{equation}
where $\enucdot$ is the specific nuclear energy generation rate and $L_R$ is the Rossby length 
described by Equation \ref{eqn:rossby}. This finding was consistent with the results of 
\citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby 
length. As seen in Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling of the 
flame speed with rotation rate in this set of runs. We observed earlier, however, that nuclear 
reactions progress more quickly at higher rotation rate. This results in a higher $\enucdot$ --- up 
to four to six times higher near the burning front after the flame ignites --- which may counteract 
the reduction in flame speed from the increased Coriolis confinement. We also observe that 
$\tt{F500}$ accelerates faster than $\tt{F1000}$, which appears to experience a small deceleration 
at early times. This disparity may be a direct result of the difference in Coriolis force.
